\bfseries\LARGE Vision Vs. LiDAR/Radar Fusion in Autonomous Driving Systems

### 2025-05-25 Papers/Paper One/TopicOne.tex
edited paper one

As vehicles become more technologically advanced, the demand for autonomous capabilities continues to rise. Features such as lane keeping, adaptive cruise control, and highway autopilot—collectively known as advanced driver-assistance systems (ADAS)—form the early building blocks of full self-driving stacks. Industry leaders now face a critical decision: how should vehicles perceive the road ahead?

@@

This report compares the two paradigms by analyzing their theoretical foundations, energy profiles, data fidelity, and robustness under real-world driving conditions. The goal is to determine whether vision-primary perception can provide sufficient safety and performance for large-scale deployment.

\section{Theory and Technical Background}

@@

The Society of Automotive Engineers (SAE) J3016 standard defines six levels of
vehicle automation \cite{SAEJ3016_2021}. These range from Level 0 (no
automation) to Level 5 (full automation without any human involvement).

@@

@@

\subsubsection{Pulsed ToF:}

\begin{equation}
R_{\text{pulse}} = \frac{c}{2} t_{\text{pulse}}
\label{eq:tof_pulse}
\end{equation}

This high-power method provides centimeter accuracy.

\subsubsection{AMCW (Amplitude-Modulated Continuous Wave):}

\begin{equation}
R_{\text{AMCW}} = \frac{c}{2} \cdot \frac{\Delta\Phi}{2\pi f_M}
\label{eq:amcw}
\end{equation}

Here, \( \Delta\Phi \) is the phase shift, and \( f_M \) is the modulation frequency \cite{2019Royo}.

\subsubsection{FMCW (Frequency-Modulated Continuous Wave):}

\begin{equation}
R_{\text{FMCW}} = f_r \cdot \frac{c T}{2B}
\label{eq:fmcw}
\end{equation}

Where \( f_r \) is the beat frequency, \( T \) the chirp duration, and \( B \) the bandwidth \cite{2019Royo}.

\subsubsection{Voxelization:}
What results from the utilization of these methods is a ``point cloud" which
corresponds to the distances the objects are away from the transmitter. The
interpretation of these distances is a process called Voxelization. There are
two forms, hard voxelization and dynamic voxelization. The in-depth explanation
of these are not required for further elaboration in this paper. Details are
shown in Figure~\ref{fig:voxelization}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{voxelization.png}
	\caption{The combining of the point cloud forms an image \autocite{RT3D}}
	\label{fig:voxelization}
\end{figure}

@@

Radar calculates velocity via Doppler frequency shift:

\begin{equation}
v = \frac{f_D \lambda}{2}
\label{eq:radar_doppler}
\end{equation}

Where \( f_D \) is the Doppler shift and \( \lambda \) is the radar wavelength \cite{Han2023FourDRadarSurvey}. Radar’s all-weather capability makes it ideal for fallback use.

\subsection{Vision-Based Depth Estimation}

Depth from stereo vision is estimated by:

\begin{equation}
D = \frac{f b}{d}
\label{eq:depth_stereo}
\end{equation}

Where \( f \) is focal length, \( b \) baseline, and \( d \) disparity.

Monocular depth uses learned inverse depth:

\begin{equation}
\hat{D}^{-1} = g(I; \theta)
\label{eq:depth_mono}
\end{equation}

Where \( g \) is a neural network with parameters \( \theta \) \cite{Li2022BEVFormer}.

\subsection{Sensor Fusion and Noise Correlation}

Noise in sensor fusion is modeled as:

\begin{equation}
\sigma_{D_f}^2 = \sum_{i=1}^{n} w_i^2\,\sigma_{D_i}^2 + 2\sum_{i<j} w_i w_j\,\rho_{ij}\,\sigma_{D_i}\sigma_{D_j}
\label{eq:noise}
\end{equation}

Here, \( w_i \) are weights, \( \sigma_{D_i} \) are variances, and \( \rho_{ij} \) are sensor correlations \cite{Rana2023PerceptionSystems}. It is clearly shown that with the inclusion of more sensors, the potential for
more noise is increased.

\subsection{Energy and Compute Cost}

Power usage scales with input bandwidth \( B \) and model complexity \( C \):

\begin{equation}
P \propto B \cdot C
\label{eq:power_model}
\end{equation}

Vision stacks operate efficiently at \SI{2.1}{\watt}, compared to \SIrange{12}{20}{\watt} for fusion \cite{Chen2024EndToEndAD, Rana2023PerceptionSystems}.
\section{System Design and Methods}

@@

Each frame, typically $128 \times 256$ pixels, is passed through a VQGAN encoder
that produces up to 512 discrete tokens per image. This reduces bandwidth while preserving high-level semantics such as lane markings and moving objects.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{architecture}
    \caption{Vision-based pipeline with VQGAN tokenization and transformer rollout\autocite{Esser2021TamingTransformersHighResolutionImage}.}
    \label{fig:tokenizer}
\end{figure}

@@

\subsubsection{Autoregressive Planning with Transformers}

The token sequence is processed by a decoder-only transformer, which predicts the next spatial token at each step. This architecture parallels natural language processing and enables generalization to unseen traffic scenarios.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{architecture2}
    \caption{Transformer model for sequence prediction from visual tokens \autocite{vaswani2023attentionneed}.}
    \label{fig:transformer}
\end{figure}

The transformer is trained using imitation learning on datasets like commaVQ and can be fine-tuned using reinforcement learning for policy robustness \autocite{goff2025learningdriveworldmodel}.

@@

\begin{enumerate}[label=\alph*), nosep]
  \item Data Acquisition: Captures and synchronizes LiDAR point clouds, radar waveforms, and camera frames.
  \item Preprocessing: Voxelizes LiDAR data, computes Doppler velocities from radar, and encodes image features.
  \item Fusion Layer: Combines sensor features using early, mid, or late neural fusion strategies.
  \item Detection and Tracking: Predicts 3D bounding boxes and tracks objects across time.
\end{enumerate}

\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{lidararchitecture}
	\caption{Reference LiDAR/radar fusion architecture \autocite{Haghighi2024}.}
	\label{fig:lidarstack}
\end{figure}

@@

\begin{itemize}[nosep]
  \item High Power Draw: LiDAR alone draws \SIrange{10}{20}{\watt}, compared to \SI{2.1}{\watt} for vision-only stacks \autocite{Chen2024EndToEndAD}.
  \item Latency: Voxelization and multi-sensor alignment add \SI{20}{\milli\second} delay, reducing real-time responsiveness \autocite{Rana2023PerceptionSystems}.
  \item Noise Correlation: Overlapping modalities propagate shared errors, increasing total uncertainty as shown in Equation~\ref{eq:noise} \autocite{Rana2023PerceptionSystems}.
  \item Maintenance Complexity: Fusion requires precise calibration and can drift over time.
\end{itemize}

This complexity restricts scalability, particularly in cost-sensitive markets.

Radar bridges the gap between vision and LiDAR. It estimates object velocity
using the Doppler effect shown in Equation~\ref{eq:radar_doppler}.

Radar is immune to adverse weather and lighting and can detect motion through occlusions. When fused with vision using a transformer-based late-fusion module, it recovers over \SI{80}{\percent} of lost mAP in foggy conditions \autocite{Liao2024RadarVisionFusion}.

A hybrid architecture combining vision and radar enables a balance of robustness and efficiency:

\begin{itemize}[nosep]
  \item Vision for primary perception—high-resolution, low-cost, and semantically rich.
  \item Radar for adverse conditions—weather-resilient and velocity-sensitive.
  \item LiDAR for benchmarking only—useful for training ground truth, not live deployment.
\end{itemize}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{multimodelarch}
	\caption{Modular architecture integrating vision, radar, and LiDAR inputs \autocite{Hasanujjaman2023}.}
	\label{fig:multiarch}
\end{figure}

This design maximizes generalization while minimizing cost and compute demands, and it reflects a biologically plausible model: humans drive with vision, but radar is the backup.

% -------------------- Results & Discussion ------------------
\section{Results and Discussion}

This section evaluates the empirical performance of vision-primary and LiDAR/radar fusion architectures. The comparison spans five axes: detection accuracy, computational efficiency, robustness, system complexity, and deployment scalability. Results are drawn from benchmarks using KITTI, nuScenes, and commaVQ datasets, as well as simulation and power profiling studies.

\subsection{Detection Accuracy and Depth Estimation}

On KITTI, the BEVFormer vision model achieved an AbsRel of 0.112 and RMSE of \SI{3.97}{\meter}, closely tracking the LiDAR-derived ground truth \autocite{Li2022BEVFormer}. On nuScenes, the vision-only configuration reached 56.2 3D mAP versus 61.5 mAP for the full fusion stack \autocite{Zhang2023MultiSensorFusionSurvey}.

However, when radar fallback was selectively integrated using transformer-based late fusion, the vision model recovered to 59.8 mAP \autocite{Liao2024RadarVisionFusion}. This narrows the performance delta to under 2 points—well within tolerances for safe operation—demonstrating that vision can match LiDAR-fusion with minimal augmentation.

\subsection{Computational Latency and Power Efficiency}

Inference benchmarks revealed stark contrasts. The vision-primary pipeline completed end-to-end perception and control in under \SI{9.8}{\milli\second}, consuming just \SI{2.1}{\watt} on an embedded platform \autocite{Chen2024EndToEndAD}.

@@

Under clear conditions, vision models outperform. In fog, heavy rain, or low-light scenarios, however, performance degraded by up to \SI{12.6}{\percent} in mAP \autocite{Han2023FourDRadarSurvey}. Transformer-based radar fallback restored more than \SI{80}{\percent} of this loss, effectively covering vision’s blind spots.

@@

Sensor fusion introduces noise amplification when data is temporally misaligned
or spatially correlated. The fusion error propagation model in
Equation~\ref{eq:noise} confirms this degradation in high-$\rho$ configurations, such as LiDAR-camera overlaps \autocite{Rana2023PerceptionSystems}. Vision stacks, trained end-to-end, maintain internal consistency and minimize accumulated error.

@@

In the biological analogy, humans navigate using vision as their primary sensor. We don’t emit lasers—we perceive, interpret, and act. This report adopts the same philosophy: with a learned representation pipeline, vision is sufficient for most environments. Radar is reserved not as a crutch, but as a safety net.

@@

\subsection{Summary}

The data is unambiguous:

\begin{itemize}
  \item Vision stacks perform within $5\%$ of LiDAR/radar fusion across key metrics, achieving 56.2 mAP on nuScenes compared to 61.5 for fusion \autocite{Zhang2023MultiSensorFusionSurvey}.
  \item With radar fallback, this gap shrinks to $<2\%$, recovering performance to 59.8 mAP \autocite{Liao2024RadarVisionFusion}.
  \item Vision systems run 3–5x faster and consume 6–10x less
		power—\SI{9.8}{\milli\second} and \SI{2.1}{\watt} vs. $\geq$ \SI{30}{\milli\second} and \SIrange{12}{20}{\watt} for fusion \autocite{Chen2024EndToEndAD, Rana2023PerceptionSystems}.
  \item Vision stacks avoid fusion noise \autocite{Rana2023PerceptionSystems}, reduce calibration burden \autocite{Han2023FourDRadarSurvey}, and cost significantly less, with LiDAR accounting for over \SI{75}{\percent} of ADS hardware expense \autocite{Shetty2022LiDARvsCamera, Sajjad2021ComparativeDetection}.
\end{itemize}

In short, vision is sufficient for autonomy. Radar makes it resilient. LiDAR
makes it expensive, but can still have its uses in training datasets as
benchmarks. The proposed ideal multi-modal solution is outlined in
Figure~\ref{fig:multimodal}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{multimodal}
	\caption{A Multi-Modal Approach to ADS \autocite{Hasanujjaman2023}}
	\label{fig:multimodal}
\end{figure}
\section{Conclusion}

This study set out to evaluate whether camera-first autonomous driving systems can match or exceed the performance of LiDAR/radar fusion approaches. The answer is clear: for most conditions, they can—and they do so with greater efficiency, lower cost, and fewer integration challenges.

The evidence shows:

\begin{itemize}
  \item Vision-primary pipelines achieve near-parity with fusion systems on 3D object detection and depth accuracy.
  \item Their energy and latency profiles outperform fusion systems by over 5x, enabling real-time inference on embedded hardware.
  \item Environmental failure modes, such as fog or night driving, can be mitigated with selective radar fallback, avoiding the need for costly LiDAR.
  \item End-to-end training on vision stacks reduces noise propagation and obviates sensor calibration.
\end{itemize}

These findings point to a clear path forward: deploy vision-first systems as the default, integrate radar for robustness where needed, and phase out LiDAR except in high-precision edge cases. With advances in simulation, world modeling, and vision transformers, the field now has the tools to deliver scalable autonomy without exotic sensors.

The next generation of autonomous vehicles will not be defined by how many sensors they carry, but by how intelligently they interpret the world with less. Vision, when empowered by the right models, is not just sufficient—it is optimal.
\footnote{LLM was used strictly for formatting IAW IEEE Editorial Style Manual
	for Authors. All thoughts, prose, and conclusions are
my own.}
\newpage
\addcontentsline{toc}{section}{References}
\printbibliography

>>>>

As vehicles become more technologically advanced, the demand for autonomous
capabilities continues to rise. Features such as lane keeping, adaptive cruise
control, and highway autopilot—collectively known as autonomous driving systems (ADS)—form the early building blocks of full self-driving stacks. Industry leaders now face a critical decision: how should vehicles perceive the road ahead?

@@

This report compares the two paradigms by analyzing their theoretical foundations, energy profiles, data fidelity, and robustness under real-world driving conditions.

\section{Technology Description}

@@

The Society of Automotive Engineers (SAE) J3016 standard defines six levels of
vehicle automation \cite{SAEJ3016_2021}. These range from Level 0 (no
automation) to Level 5 (full automation without any human involvement). The full
description of each level is given in Table~\ref{tab:sae}.

@@

Currently, most ADS operate at Level 2, which is very much like being a
driver's ed instructor sitting at secondary controls while a 15-year-old is
operating the vehicle, ready to take control at any given moment. Level 5 is
ideally like sitting in a train car, with
absolutely no attention needed from the driver. This frees the driver to work on
other important tasks such as grading papers, concentrating on a new idea, or
taking some real downtime at the end of a stressful day.

@@

\subsubsection{Pulsed ToF:}
Pulsed TOF techniques are the simplest of the three cases in which distance is
determined by multiplying the speed of light in a medium by the time a light
pulse takes to travel the distance to the target:
\begin{equation}
R_{\text{pulse}} = \frac{c}{2} t_{\text{oF}},
\label{eq:tof_pulse}
\end{equation}
where $R_{oF}$ is the distance the light ray travels, $c$ is the speed of
light in free space, and $t_{oF}$ is the time it takes for the pulse of energy
to travel from its emitter to the observed object and then back to the receiver.
This high-power method provides centimeter accuracy. This method is shown in
Figure~\ref{fig:pulse}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{lidar.png}
	\caption{The Pulsed Approach to Determining Object Distance \cite{2019Royo}}
	\label{fig:pulse}
\end{figure}

\subsubsection{AMCW (Amplitude-Modulated Continuous Wave):}
The AMCW approach modulates the intensity of a continuous lightwave instead of
laser pulsing as utilized in Eqn.~\ref{eq:tof_pulse}. This method utilizes the
corresponding shift in phase in the received signal to calculate the distance to
the object, $R$:
\begin{equation}
R = \frac{c}{2} \cdot \frac{\Delta\Phi}{2\pi f_M},
\label{eq:amcw}
\end{equation}
where \( \Delta\Phi \) is the phase shift, $c$ is the speed of light in free
space, and \( f_M \) is the modulation
frequency \cite{2019Royo}. It's simplified principles are shown in
Figure~\ref{fig:amcw}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{amcw}
	\caption{Measuring Distance by Phase Difference \cite{2019Royo}}
	\label{fig:amcw}
\end{figure}
\subsubsection{FMCW (Frequency-Modulated Continuous Wave):}
This final method involves a modulating the power applied to the source
\cite{petermannopto}. The signal returned is mixed with the source signal,
creating a beat frequency that is a measure of the probed distance, $R$:
\begin{equation}
R = f_r\frac{c T}{2B},
\label{eq:fmcw}
\end{equation}
where \( f_r \) is the beat frequency, \( T \) the period of the ramp, $c$ is
the speed of light in free space, and \( B \)
the bandwidth of the frequency sweep \cite{2019Royo}. The main parameters are
shown in Figure~\ref{fig:fmcw}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{fmcw}
	\caption{Measuring Distance by Frequency Modulation \cite{2019Royo}}
	\label{fig:fmcw}
\end{figure}
This method is fundamentally different as there is additional processing
overhead done in the Fourier domain rather than in the time or complex domain
for the pulsed and AMCW approach respectively. This ultimately allows precision
to much higher precision \cite{petermannopto}.

\subsubsection{Voxelization:}
What results from the utilization of these methods is a ``point cloud" which
corresponds to the distances the objects are away from the transmitter. The
interpretation of these distances into 3D object boxes is a process called
Voxelization. The high level process of breaking down point clouds into voxels and
interpreting them is shown in Figure~\ref{fig:voxelization}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\textwidth]{voxelization.png}
	\caption{The combining of the point cloud forms an image where (a) is the
	interpreted collection of point clouds inferred by the distance calculation in the sensor, (b) is the
segmentation of this set of distances into a grid to create a 'set of sets' of
different points, (c) is the transformer pipeline which leads to object
identification, and (d) is the outcome with identified objects boxed
\autocite{RT3D}.}

	\label{fig:voxelization}
\end{figure}

@@

The Doppler effect is the change in frequency or wavelength of a wave for an
observer or sensor, who is moving relative to the wave source. Radar calculates velocity $v$, via Doppler frequency shift:

\begin{equation}
v = \frac{f_D \lambda}{2},
\label{eq:radar_doppler}
\end{equation}
where \( f_D \) is the Doppler shift and \( \lambda \) is the radar wavelength
\cite{Han2023FourDRadarSurvey}. Radar’s all-weather capability makes it ideal
for fallback use. Figure~\ref{fig:doppler} shows the Doppler effect as it
pertains to humans, a useful tool in understanding how this phenomena can be
applied to this.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.4\textwidth]{doppler}
	\caption{The classic Doppler example showing an ambulance emitting sound. An
	observer, from whom the ambulance leaving from will receive fewer pressure and
will hear a deeper note, while an observer from whom the ambulance is
approaching will receive higher pressure fluctuations and hear a higher note.}
	\label{fig:doppler}
\end{figure}
\subsection{Sensor Fusion and Noise Correlation}

Noise is a deviation of the response to the ideal signal and manifests itself
differently depending on the type of sensor in phenomena such as temperature,
EMF, or visual interference (weather). Noise in sensor fusion is modeled as:
\begin{equation}
\sigma_{D_f}^2 = \sum_{i=1}^{n} w_i^2\,\sigma_{D_i}^2 + 2\sum_{i<j} w_i
w_j\,\rho_{ij}\,\sigma_{D_i}\sigma_{D_j},
\label{eq:noise}
\end{equation}

where \( w_i \) are weights, \( \sigma_{D_i} \) are variances, and \( \rho_{ij}
\) are sensor correlations \cite{Rana2023PerceptionSystems}.

@@

Each frame, typically $128 \times 256$ pixels, is passed through a VQGAN encoder
that produces up to 512 discrete tokens per image. This reduces bandwidth while
preserving high-level semantics such as lane markings and moving objects. The
process of encoding these images into tokens is shown in
Figure~\ref{fig:tokenizer}.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{architecture}
    \caption{Vision-based pipeline with VQGAN tokenization and transformer rollout \autocite{Esser2021TamingTransformersHighResolutionImage}.}
    \label{fig:tokenizer}
\end{figure}

@@

\subsubsection{Autoregressive Planning with Transformers}
Much like how an LLM can predict the next word in a string through tokenization,
a transformer can predict a variety of outcomes in a driving model from a static
image that is tokenized. The token sequence is processed by a decoder-only
transformer, which predicts the next spatial token at each step. This
architecture parallels natural language processing and enables generalization to
unseen traffic scenarios \cite{drivingsim}. The modern transformer architecture
employed, shown in Figure~\ref{fig:transformer}, is taken from one of the most
important works in recent AI advancement, ``Attention is All You Need"
\cite{vaswani2023attentionneed}.
\begin{figure}[H]
    \centering
    \includegraphics[width=0.5\textwidth]{architecture2}
    \caption{Transformer model for sequence prediction from visual tokens \autocite{vaswani2023attentionneed}.}
    \label{fig:transformer}
\end{figure}

The transformer is trained using imitation learning on datasets like commaVQ and
can be fine-tuned using reinforcement learning for policy robustness
\autocite{drivingsim}. This way, the training data need not have to contain
every situation imaginable, as the process itself will be able to infer all
possible world conditions \cite{goff2025learningdriveworldmodel}. Once all of
these policies are applied and processing is applied, the details in
Figure~\ref{fig:visionpipe} can be inferred.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{masking}
	\caption{The Vision Pipeline shows the original camera image, the feature map,
	and the image mask \cite{wang2024}.}
	\label{fig:visionpipe}
\end{figure}

@@

\begin{enumerate}[label=\alph*), nosep]
  \item Data Acquisition: Captures and synchronizes LiDAR point clouds, radar waveforms, and camera frames.
  \item Preprocessing: Voxelizes LiDAR data, computes Doppler velocities from radar, and encodes image features.
  \item Fusion Layer: Combines sensor features using early, mid, or late neural fusion strategies.
  \item Detection and Tracking: Predicts 3D bounding boxes and tracks objects across time.
\end{enumerate}
Figure~\ref{fig:lidarstack} shows how
Eqns.~\ref{eq:tof_pulse},~\ref{eq:amcw}, and ~\ref{fig:fmcw} can be utilized and
combined to overcome noise and predict 3D bounding boxes.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{lidararchitecture}
	\caption{Reference LiDAR/radar fusion architecture \autocite{Haghighi2024}.}
	\label{fig:lidarstack}
\end{figure}

@@

\begin{itemize}[nosep]
  \item High Power Draw: LiDAR alone draws \SIrange{10}{20}{\watt}, compared to \SI{2.1}{\watt} for vision-only stacks \autocite{Chen2024EndToEndAD}.
  \item Latency: Voxelization and multi-sensor alignment add \SI{20}{\milli\second} delay, reducing real-time responsiveness \autocite{Rana2023PerceptionSystems}.
  \item Noise Correlation: Overlapping modalities propagate shared errors,
		increasing total uncertainty as shown in Eqn.~\ref{eq:noise} \autocite{Rana2023PerceptionSystems}.
  \item Maintenance Complexity: Fusion requires precise calibration and can drift over time.
\end{itemize}
\begin{figure}[H]
	\centering
	\includegraphics[width=\textwidth]{voxels}
	\caption{Representations of point clouds from LiDAR/Radar are raw point
		clouds, post voxel-sampling, and the generated the point clouds view in
	Birds-Eye-View \cite{wang2024}.}
\label{fig:voxels}
\end{figure}

This complexity restricts scalability, particularly in cost-sensitive markets.
Figure~\ref{fig:voxels} shows a typical LiDAR/Radar pipeline \cite{wang2024}.

% -------------------- Results & Discussion ------------------
\section{Comparisons}

This section evaluates the empirical performance of vision-primary and
LiDAR/Radar fusion architectures. The comparison spans five axes: detection
accuracy, computational efficiency, robustness, system complexity, and
deployment scalability. Results are drawn from benchmarks using
KITTI \autocite{kitti}, nuScenes \autocite{nuscenes}, and commaVQ
\autocite{goff2025learningdriveworldmodel} datasets, as well as simulation and
power profiling studies \autocite{Rana2023PerceptionSystems}.

\subsection{Detection Accuracy and Depth Estimation}
In order to properly classify and compare these technologies, a method to
evaluate them must be first established. For this, \textit{Precision} is the
measure of the probability of a classifier predicting a true example and
\textit{recall} is the number of positive samples correctly identified versus
the number of images whose class is really a positive class, expressed by:
\begin{equation}
	recall = \frac{TP}{TP + FN},
	\label{eq:recall}
\end{equation}
where $TP$ is the number of true positives, $FP$ is the number of false
positives, and $FN$ is the number of false negatives.

Building off of this, the most commonly used evaluation metric for object
detection is $AP$, or Average Precision. This is defined by:
\begin{equation}
	AP = \int_0^1 p(r)dr,
	\label{eq:ap}
\end{equation}
where $p(r)$ is the function of precision about the recall.

However, for a
better statistical representation of the data, a more popular metric is used
defined as $mAP$, or Mean Average Position, which can be expressed as:
\begin{equation}
	mAP = \frac{\sum_{i=1}^N AP_i}{N},
	\label{eq:map}
\end{equation}
where, $N$ is the total number of classes and $AP_i$ denotes the average
precision of the $i$-th class \cite{wang2024}.

\subsection{Findings}
On KITTI, the BEVFormer vision model achieved an AbsRel of 0.112 and RMSE of
\SI{3.97}{\meter}, closely tracking the LiDAR-derived ground truth
\autocite{Li2022BEVFormer}. On nuScenes, the vision-only configuration reached
56.2 3D $mAP$ versus 61.5 $mAP$ for the full fusion stack \autocite{Zhang2023MultiSensorFusionSurvey}.

However, when radar fallback was selectively integrated using transformer-based
late fusion, the vision model recovered to 59.8 $mAP$ \autocite{Liao2024RadarVisionFusion}. This narrows the performance delta to under 2 points—well within tolerances for safe operation—demonstrating that vision can match LiDAR-fusion with minimal augmentation.

\subsection{Computational Latency and Power Efficiency}
Inference benchmarks revealed stark contrasts. The vision-primary pipeline completed end-to-end perception and control in under \SI{9.8}{\milli\second}, consuming just \SI{2.1}{\watt} on an embedded platform \autocite{Chen2024EndToEndAD}.

@@

Under clear conditions, vision models outperform. In fog, heavy rain, or
low-light scenarios, however, performance degraded by up to \SI{12.6}{\percent}
in $mAP$ \autocite{Han2023FourDRadarSurvey}. Transformer-based radar fallback restored more than \SI{80}{\percent} of this loss, effectively covering vision’s blind spots.

@@

Sensor fusion introduces noise amplification when data is temporally misaligned
or spatially correlated. The fusion error propagation model in
Eqn.~\ref{eq:noise} confirms this degradation in high-$\rho$ configurations,
such as LiDAR-camera overlaps \autocite{Rana2023PerceptionSystems}. Vision
stacks, trained end-to-end, maintain internal consistency and minimize
accumulated error \cite{goff2025learningdriveworldmodel}.

@@

Elon Musk famously claimed ``LiDAR is a fool's errand" to much surprise at the
time \cite{TeslaAutonomyDay2019}. In 2019, this was perhaps too forward
thinking, but was it without bounds? In the biological analogy, humans navigate using vision as their primary sensor. We don’t emit lasers—we perceive, interpret, and act.

@@

Radar bridges the gap between vision and LiDAR. It estimates object velocity
using the Doppler effect shown in Eqn.~\ref{eq:radar_doppler} and in
Figure~\ref{fig:doppler}.

Radar is immune to adverse weather and lighting and can detect motion through
occlusions. When fused with vision using a transformer-based late-fusion module,
it recovers over \SI{80}{\percent} of lost $mAP$ in foggy conditions
\autocite{Liao2024RadarVisionFusion}. In low speed, high traffic conditions,
radar can be utilized to better estimate distance to other vehicles with less
cooling necessary. This is very enticing for high population areas such as Los
Angeles where stop and go traffic on exceedingly hot roads is a daily occurrence.

What companies such as Comma.ai and Tesla are doing is employing a hybrid
architecture combining vision and radar which enable a balance of robustness and efficiency:

\begin{itemize}[nosep]
  \item Vision for primary perception—high-resolution, low-cost, and semantically rich.
  \item Radar for adverse conditions—weather-resilient and velocity-sensitive.
  \item LiDAR for benchmarking only—useful for training ground truth, not live deployment.
\end{itemize}
Figure~\ref{fig:multiarch} shows the modular stack which can be tuned via
hyperparameters during the training stage to give more weight to favor certain
inputs. Such an approach is utilized by Comma.ai to prioritize key sensor inputs
to the model deployed in Openpilot \cite{goff2025learningdriveworldmodel}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.65\textwidth]{multimodelarch}
	\caption{Modular architecture integrating vision, radar, and LiDAR inputs \autocite{Hasanujjaman2023}.}
	\label{fig:multiarch}
\end{figure}

This design maximizes generalization while minimizing cost and computation demands, and it reflects a biologically plausible model: humans drive with vision, but radar is the backup.

\subsection{Summary}
In order to eventually achieve SAE Level 5 Autonomy, the currently existing data points to the possibility of vision-only stacks
surpassing fusion models in the near future:
\begin{itemize}
  \item Vision stacks perform within $5\%$ of LiDAR/Radar fusion across key
		metrics, achieving 56.2 $mAP$ on nuScenes compared to 61.5 for fusion \autocite{Zhang2023MultiSensorFusionSurvey}.
  \item With radar fallback, this gap shrinks to $<2\%$, recovering performance
		to 59.8 $mAP$ \autocite{Liao2024RadarVisionFusion}.
  \item Vision systems run 3–5x faster and consume 6–10x less
		power—\SI{9.8}{\milli\second} and \SI{2.1}{\watt} vs. $\geq$ \SI{30}{\milli\second} and \SIrange{12}{20}{\watt} for fusion \autocite{Chen2024EndToEndAD, Rana2023PerceptionSystems}.
  \item Vision stacks avoid fusion noise \autocite{Rana2023PerceptionSystems}, reduce calibration burden \autocite{Han2023FourDRadarSurvey}, and cost significantly less, with LiDAR accounting for over \SI{75}{\percent} of ADS hardware expense \autocite{Shetty2022LiDARvsCamera, Sajjad2021ComparativeDetection}.
\end{itemize}

In short, vision is sufficient for autonomy. Radar makes it resilient. LiDAR
makes it expensive, but can still have its uses in training datasets as
benchmarks. The ideal multi-modal solution is outlined in
Figure~\ref{fig:multimodal}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.7\textwidth]{multimodal}
	\caption{A Multi-Modal Approach to ADS \autocite{Hasanujjaman2023}}
	\label{fig:multimodal}
\end{figure}
\section{Conclusion}

This paper sets out to compare the existing technology employed in Level 2 ADS systems
deployed in production around the world today. While approaches such as Waymo's
LiDAR/Vision/Radar fusion are commendable and show great promise in deployment, the
drawbacks mentioned and advancements in the vision stack in particular offer a
clear alternative that is cheaper and more energy efficient becoming more effective each day.
\footnote{An LLM was used strictly for formatting IAW IEEE Editorial Style Manual
	for Authors. All thoughts, prose, and conclusions are
my own.}
\newpage
\addcontentsline{toc}{section}{References}
\printbibliography
